Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a  Mesoscale Convective System Case During RELAMPAGO campaign

### 2022-08-09 analysis/paper/paper_ana_20181122.Rmd
Agrega algunas figuras

layout: authoryear,preprint,review,12pt
#layout: final,5p,times,twocolumn,authoryear

@@

```{r case, fig.cap="(ref:case)"}

@@

  .[time != ymd_h(2018112312)] %>% 
  .[, ":="(espesor = (`500`-`1000`)/100, time = factor(time))] %>%
  ggplot(aes(longitude, latitude)) +
  geom_contour_fill(aes(z = cape, fill = stat(level_d)), breaks = seq(500, 4000, 500)) +
  scale_fill_distiller(super = ScaleDiscretised,
                       name = "CAPE",
                       # palette = "PRGn", direction = 1,
                       palette = "YlOrRd", direction = 1,
                       guide = guide_colorsteps(barwidth = 25, mid = 25,
                                                barheight = 0.5,
                                                title.position = "left",
                                                title.vjust = 1,
                                                frame.colour = "black")) +
  geom_mapa() +
  geom_contour2(aes(z = espesor, label = ..level..),
                size = 0.2, color = "darkred", linetype = 2,
                label_size = 2, label.placer = label_placer_n(n = 3)) +
  # label_size = 2, label.placer = label_placer_fraction(frac = 0.75)) +
  geom_contour2(aes(z = msl/100, label = ..level..),
                size = 0.4, color = "black", linetype = 1,
                label_size = 2) +
  cordillera +
  coord_sf(ylim = c(-45, -22), xlim = c(-75, -52)) +
  facet_wrap(vars(time), labeller = labeller(time = c("2018-11-22 00:00:00" = "00 UTC Nov 22",
                                                      "2018-11-22 12:00:00" = "12 UTC Nov 22",
                                                      "2018-11-23 00:00:00" = "00 UTC Nov 23"))) +
  tag_facets() +
  labs(x = NULL, y = NULL, fill = "CAPE" ) +
  theme_minimal() +
  theme(legend.position = "bottom")

@@

We used multi-physics scheme to better represent the uncertainty in the model formulation within the DA system. We generated 9 different model configurations consisting of the combination of 3 moist convection schemes (Kain–Fritsch [@kain2004], Grell–Freitas [@grell2013], and Betts–Miller–Janjic [@janjic1994]) and 3 planetary boundary layer schemes (Yonsei University Scheme [@hong2006], Mellor–Yamada–Janjic Scheme [@janjic1994], and Mellor–Yamada Nakanishi Niino [@nakanishi2009]). All ensemble members use the same land-surface model (Noah-MP, @chen2001), microphysics (WRF single-moment 6–class scheme [@hong2006a]), and radiation processes (RRTMG shortwave and longwave scheme [@iacono2008]) parameterizations.

@@

(ref:UV-diff) Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS, and c) and f) RAD-SATWND for the spatially averaged vertical profiles of u wind (a, b, and c, in $m\ s^{-1}$) and v wind (d, e, and f in $m\ s^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle. Black contours correspond to u wind and aashed contours to negative v wind for (a) AWS, (b) SATWND, and (c) RAD and v wind for (d) AWS, (e) SATWND, and (f) RAD since those experiments are the ones with more assimilated observations in each panel.

@@

@@

  ana[bottom_top %between% c(1, 10), .(theta_mean = mean(theta)), 
      by = .(south_north, west_east, date, exp)] %>% 
  .[pw, on = .NATURAL] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = theta_mean, fill = stat(level_d)),
                    proj = norargentina_lambert,
                    breaks = c(-Inf, seq(290, 314, 2), Inf), limits = c(-Inf, Inf)) +
  scale_fill_distiller(super = ScaleDiscretised,
                       palette = "RdYlBu", direction = -1,
                       guide = guide_colorsteps(barwidth = 25,
                                                barheight = 0.5, 
                                                title.position = "left", 
                                                title.vjust = 1,
                                                frame.colour = "black")) +
  labs(fill = "Potential\ntemperature (K)") +
  cordillera +
  geom_mapa() +
  facet_grid(. ~ exp, 
             labeller = labeller(exp = c(E2 = "CONV", E5 = "AWS", 
                                         E6 = "SATWND", E9 = "RAD"))) +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3),
        strip.text = element_blank(), 
        strip.background = element_blank())

>>>>

# layout: authoryear,preprint,review,12pt
layout: final,5p,times,twocolumn,authoryear

@@

```{r case, fig.cap="(ref:case)", fig.width=7, fig.height=3.5, fig.align="center", fig.env = "figure*"}

@@

  .[time != ymd_h(2018112312)] %>% 
  .[, ":="(espesor = (`500`-`1000`)/100, time = factor(time))] %>%
  ggplot(aes(longitude, latitude)) +
  geom_contour_fill(aes(z = cape, fill = stat(level_d)), breaks = seq(500, 4000, 500)) +
  scale_fill_distiller(super = ScaleDiscretised,
                       name = "CAPE",
                       # palette = "PRGn", direction = 1,
                       palette = "YlOrRd", direction = 1,
                       guide = guide_colorsteps(barwidth = 25, mid = 25,
                                                barheight = 0.5,
                                                title.position = "left",
                                                title.vjust = 1,
                                                frame.colour = "black")) +
  geom_mapa() +
  geom_contour2(aes(z = espesor, label = ..level..),
                size = 0.2, color = "darkred", linetype = 2,
                label_size = 2, label.placer = label_placer_n(n = 3)) +
  # label_size = 2, label.placer = label_placer_fraction(frac = 0.75)) +
  geom_contour2(aes(z = msl/100, label = ..level..),
                size = 0.4, color = "black", linetype = 1,
                label_size = 2) +
  cordillera +
  coord_sf(ylim = c(-45, -22), xlim = c(-75, -52)) +
  facet_wrap(vars(time), labeller = labeller(time = c("2018-11-22 00:00:00" = "00 UTC Nov 22",
                                                      "2018-11-22 12:00:00" = "12 UTC Nov 22",
                                                      "2018-11-23 00:00:00" = "00 UTC Nov 23"))) +
  tag_facets() +
  labs(x = NULL, y = NULL, fill = "CAPE" ) +
  theme_minimal(base_size = 8) +
  theme(legend.position = "bottom")

@@

We used multi-physics scheme to better represent the uncertainty in the model formulation within the DA system. We generated 9 different model configurations consisting of the combination of 3 moist convection schemes (Kain–Fritsch [@kain2004], Grell–Freitas [@grell2013], and Betts–Miller–Janjic [@janjic1994]) and 3 planetary boundary layer schemes (Yonsei University Scheme [@hong2006], Mellor–Yamada–Janjic Scheme [@janjic1994], and Mellor–Yamada Nakanishi Niino [@nakanishi2009]). The distribution of these schemes among the 60 ensemble members is outlined in Table \@ref(tab:miembros-desc). The multi-physics approach is also introduced in order to represent the uncertainty associated with the more relevamnt physical processes that are not resolved by the model. All ensemble members use the same land-surface model (Noah-MP, @chen2001), microphysics (WRF single-moment 6–class scheme [@hong2006a]), and radiation processes (RRTMG shortwave and longwave scheme [@iacono2008]) parameterizations.

```{r miembros-desc}
fisica %>% 
  copy() %>% 
  .[, c("Cumulus", "PBL") := tstrsplit(fisica, split = "-")] %>% 
  .[, .(mem = paste(as.numeric(mem), collapse = ", ")), by = .(PBL, Cumulus)] %>% 
  dcast(Cumulus ~ PBL, value.var  = "mem") %>% 
  kable(booktabs = TRUE, caption = "Generation of the 60-member RRR multi-physics ensemble as a combination of Cumulus and PBL parameterizations, and GEFS members for boundary conditions.", 
        align = "cccc") %>% 
  add_header_above(c(" " = 1, PBL = 3)) %>% 
  kable_classic_2(full_width = FALSE) %>% 
  kable_styling(font_size = 6,
                position = "center") %>% 
  # column_spec(1, width = "1em") %>% 
  column_spec(2:4, width = "7em")
```

@@

(ref:UV-diff) Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS, and c) and f) RAD-SATWND for the spatially averaged vertical profiles of u wind (a, b, and c, in $m\ s^{-1}$) and v wind (d, e, and f in $m\ s^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle. Black contours correspond to u wind and dashed contours to negative v wind for (a) AWS, (b) SATWND, and (c) RAD and v wind for (d) AWS, (e) SATWND, and (f) RAD since those experiments are the ones with more assimilated observations in each panel.

@@

(ref:era5-cap) Difference between analysis ensemble mean experiments and ERA5 for the spatially averaged vertical profiles of air temperature (K, a--d), specific humidity ($g\ Kg^{-1}$, e--h) and meridional wind ($m\ s^{-1}$, i--l) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.

```{r era5, fig.cap="(ref:era5-cap)", fig.env="figure*", fig.height=7, fig.align="center"}
era <- fread("~/datosmunin3/DATA_EXP/ERA5/perfiles_ana_era5.csv")

files <- Sys.glob(here::here("/home/paola.corrales/datosmunin3/EXP//derived_data/perfiles_ana_E[2569]_new.csv"))

exp <- purrr::map(files, function(f) {
  exp <- unglue(basename(f), "perfiles_{run}_{exp}_new.csv")
  fread(f) %>% 
    .[, exp := exp[[1]][["exp"]]]

era[exp, on = c("lev", "date")] %>% 
  .[lev != 1000] %>%
  ggplot(aes(date, lev)) +
  geom_contour_fill(aes(z = T - i.T, fill = stat(level)), 
                    breaks = seq(-3, 3, 0.5)) +
  geom_contour2(aes(z = T - i.T), color = "white", size = 0.1,
                breaks = seq(-3, 3, 0.5)) +
  scale_fill_divergent_discretised(guide = guide_colorsteps(barwidth = 25,
                                                            barheight = 0.3,
                                                            title.position = "left", 
                                                            title.vjust = 1,
                                                            frame.colour = "black")) +
  scale_y_level(name = "Pressure (hPa)", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
  scale_date(ini = 20181121000000, break_bin = "12 hours") +
  facet_wrap(vars(exp), ncol = 4, labeller = labeller(exp = c(E2 = "CONV", E5 = "AWS", 
                                                              E6 = "SATWND", E9 = "RAD"))) +
  labs(fill = "Temperature (K)",
       x = NULL,
       y = "Pressure (hPa)") +
  tag_facets(position = list(x = 0.1, y = 0.96)) +
  theme_minimal(base_size = 8) +
  theme(legend.position = "bottom",
        tagger.panel.tag.text = element_text(size = 8),
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3)) +

era[exp, on = c("lev", "date")] %>% 
  .[lev != 1000] %>%
  ggplot(aes(date, lev)) +
  geom_contour_fill(aes(z = QVAPOR - i.QVAPOR, fill = stat(level)), 
                    breaks = seq(-0.0015, 0.0025, 0.0004)) +
  geom_contour2(aes(z = QVAPOR - i.QVAPOR), color = "white", size = 0.1,
                breaks = seq(-0.0015, 0.0025, 0.0004)) +
  scale_fill_divergent_discretised(guide = guide_colorsteps(barwidth = 25,
                                                            barheight = 0.3,
                                                            title.position = "left", 
                                                            title.vjust = 1,
                                                            frame.colour = "black")) +
  scale_y_level(name = "Pressure (hPa)", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
  scale_date(ini = 20181121000000, break_bin = "12 hours") +
  facet_wrap(vars(exp), ncol = 4, labeller = labeller(exp = c(E2 = "CONV", E5 = "AWS", 
                                                              E6 = "SATWND", E9 = "RAD"))) +
  labs(fill = latex2exp::TeX("Specific \nhumidity ($g$ $Kg^{-1}$)"),
       x = NULL,
       y = "Pressure (hPa)") +
  tag_facets(position = list(x = 0.1, y = 0.96), tag_pool = c("e", "f", "g", "h")) +
  theme_minimal(base_size = 8) +
  theme(legend.position = "bottom",
        tagger.panel.tag.text = element_text(size = 8),
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3)) +

  era[exp, on = c("lev", "date")] %>% 
    .[lev != 1000] %>%
    ggplot(aes(date, lev)) +
    geom_contour_fill(aes(z = V - i.V, fill = stat(level)), 
                      breaks = seq(-7, 4, 1)) +
    geom_contour2(aes(z = V - i.V), color = "white", size = 0.1,
                  breaks = seq(-7, 4, 1)) +
    scale_fill_divergent_discretised(guide = guide_colorsteps(barwidth = 25,
                                                              barheight = 0.3,
                                                              title.position = "left", 
                                                              title.vjust = 1,
                                                              frame.colour = "black")) +
    scale_y_level(name = "Pressure (hPa)", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
    scale_date(ini = 20181121000000, break_bin = "12 hours") +
    facet_wrap(vars(exp), ncol = 4, labeller = labeller(exp = c(E2 = "CONV", E5 = "AWS", 
                                                                E6 = "SATWND", E9 = "RAD"))) +
    labs(fill = latex2exp::TeX("V ($m$ $s^{-1}$)"),
         x = NULL,
         y = "Pressure (hPa)") +
    tag_facets(position = list(x = 0.1, y = 0.96), tag_pool = c("i", "j", "k", "l")) +
    theme_minimal(base_size = 8) +
    theme(legend.position = "bottom",
          tagger.panel.tag.text = element_text(size = 8),
          panel.ontop = TRUE,
          panel.grid = element_line(linetype = 3)) +

  plot_layout(ncol = 1, heights = c(1, 1, 1)) 
```

@@

  ana[bottom_top %between% c(1, 10), .(theta_mean = mean(theta)), 
      by = .(south_north, west_east, date, exp)] %>% 
  .[pw, on = .NATURAL] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = theta_mean, fill = stat(level_d)),
                    proj = norargentina_lambert,
                    breaks = c(-Inf, seq(290, 314, 2), Inf), limits = c(-Inf, Inf)) +
  scale_fill_distiller(super = ScaleDiscretised,
                       palette = "RdYlBu", direction = -1,
                       guide = guide_colorsteps(barwidth = 25,
                                                barheight = 0.5, 
                                                title.position = "left", 
                                                title.vjust = 1,
                                                frame.colour = "black")) +
  labs(fill = "Potential\ntemperature (K)") +
  cordillera +
  geom_mapa() +
  facet_grid(. ~ exp, 
             labeller = labeller(exp = c(E2 = "CONV", E5 = "AWS", 
                                         E6 = "SATWND", E9 = "RAD"))) +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3),
        strip.text = element_blank(), 
        strip.background = element_blank()) +
